Appendix

### 2024-09-06 manuscript/manuscript.Rmd
Added new plot.
Updated appendix E.

```
We then consider the empirical probability of type I errors: the probability of incorrectly classifying a centrally normal distribution as being non-normal based on the test statistic $\tau_{\text{ecn}}$. Under the assumption that the asymmetric generalised normal distributions are centrally normal after robust transformation, this produces the relationship shown in Figure \ref{fig:empirical-central-normality-type-1-error-rate}. Since for $\kappa \leq 0.80$ error curves for both transformation methods are similar, we fix the value of the central portion $\kappa$ to 0.80. In the remaining we will use the empirical central normality test statistic values defined using robust location- and scale-invariant Yeo-Johnson transformations, as it is slightly more strict. This leads to test statistic values listed in Table \ref{tab:empirical-central-normality}

@@

```{=latex}
\begin{table}
\begin{center}
\caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate.}
\label{tab:empirical-central-normality}
\begin{tabular}{c c c c c c c}

\toprule
0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\

\midrule
0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

@@

This study had the following limitations.
First, in our experimental results we focussed on the Yeo-Johnson transformation because of its ability to handle features with negative values, and did not evaluate the Box-Cox transformation.
Likewise, we focussed on MLE to optimise parameters, and did not investigate other metrics to the same extent, due to lack of demonstrable benefit (Appendix B).

>>>>

```
We then consider the empirical probability of type I errors: the probability of incorrectly classifying a centrally normal distribution as being non-normal based on the test statistic $\tau_{\text{ecn}}$.
Under the assumption that the asymmetric generalised normal distributions are centrally normal after robust transformation, this produces the relationship shown in Figure \ref{fig:empirical-central-normality-type-1-error-rate}.
Since for $\kappa \leq 0.80$ error curves for both transformation methods are similar, we fix the value of the central portion $\kappa$ to 0.80.
In the remaining we will use the empirical central normality test statistic values defined using robust location- and scale-invariant Yeo-Johnson transformations, as it is more conservative (see Appendix E).
This leads to test statistic values listed in Table \ref{tab:empirical-central-normality}

@@

```{=latex}
\begin{table}
\begin{center}
\caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate.}
\label{tab:empirical-central-normality}
\begin{tabular}{l | c c c c c c c}

\toprule
type I error rate & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\

\midrule
$\tau_{\text{ecn}}$ & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

@@

The empirical test for central normality differs from other tests for normality in that the test statistic is independent of the number of samples.
This makes this test more practical for assessing central normality for larger sample numbers, where other test may detect trivial deviations from normality.

This study had the following limitations.
First, in our experimental results we focussed on the Yeo-Johnson transformation because of its ability to handle features with negative values, and did not evaluate the Box-Cox transformation.
Likewise, we focussed on MLE to optimise parameters, and did not investigate other metrics to the same extent, due to lack of demonstrable benefit (Appendix B-D).

### 2024-09-06 manuscript/manuscript_appendix.Rmd
Added new plot.
Updated appendix E.

# Appendix D: Empirical central normality test

@@

```{r empirical-central-normality-test-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
cap <- paste0(
  "Blabla"
)

# Appendix E: Experimental results using location- and scale-invariant Box-Cox transformation

>>>>

# Appendix D: Experimental results using location- and scale-invariant Box-Cox transformation

# Appendix E: Empirical central normality test

@@

```{r empirical-central-normality-test-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
cap <- paste0(
  "Type 1 error rate as function of the test statistic $\\tau_{\\text{ecn}}$ ",
  "for five datasets, with central portion $\\kappa=0.80$. ",
  "The type 1 error rate is computed from $m_d=10000$ randomly sampled features, ",
  "These are sampled from asymmetric generalized normal distributions, with and without outliers (Box-Cox and Yeo-Johnson), ",
  "or normal distributions with outliers (none). ",
  "The test statistic is computed as the average residual of each feature after (Box-Cox and Yeo-Johnson) robust location- and shift-invariant power transformation, or before (none)."
)

data <- .plot_type_1_error_rate_appendix(plot_theme = plot_theme, manuscript_dir = manuscript_dir)
data$p
```

```{=latex}
\begin{table}
\begin{center}
\caption{Test statistic $\tau_{\text{ecn}}$ for empirical central normality at $\kappa = 0.80$ as a function of Type I error rate for several datasets.}
\label{tab:empirical-central-normality-appendix}
\begin{tabular}{l | c c c c c c c}

\toprule
type I error rate & 0.50 & 0.20 & 0.10 & 0.05 & 0.02 & 0.01 & 0.001 \\

\midrule
Box-Cox                  & 0.047 & 0.073 & 0.090 & 0.106 & 0.126 & 0.140 & 0.188 \\
Box-Cox (no outlier)     & 0.043 & 0.065 & 0.079 & 0.092 & 0.106 & 0.116 & 0.155 \\
Yeo-Johnson              & 0.041 & 0.062 & 0.075 & 0.088 & 0.103 & 0.115 & 0.154 \\
Yeo-Johnson (no outlier) & 0.041 & 0.061 & 0.074 & 0.085 & 0.099 & 0.109 & 0.139 \\
normal distr.            & 0.039 & 0.066 & 0.083 & 0.097 & 0.117 & 0.132 & 0.174 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

The results are shown in Figure \ref{fig:empirical-central-normality-test-appendix} and Table \ref {tab:empirical-central-normality-appendix}.
These indicate that the test behaves similarly for the different datasets.
For low type 1 error rates the test statistic proposed in the main manuscript is more conservative than alternatives 
based on residuals after Box-Cox transformations of asymmetric generalised normally distributed features or on residuals from strictly normally distributed features.
